---
title: "figure7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse) 
library(ggpubr) 
library(reshape2)
library(vioplot)
library(ggExtra)
library(ggplot2)
library(ggsankey)
library(paletteer)
library(survival)
library(survminer)
library(survivalROC)
```



```{r}
EGFR_CNV_riskscore <- read.csv("CNV_riskscore.csv", header=T, sep=",", check.names=F, row.names = 1, fill = TRUE)
my_comparisons <- list( c("amplification", "normal")) 

```


## Analyzing CCL2 mRNA levels in TCGA patients with or without EGFR amplification
```{r cnvEGFR-CCL2}
ggviolin(EGFR_CNV_riskscore, x = 'cnvlevel', y = 'CCL2', fill = 'cnvlevel',
         palette = c("#8785B2FF","#D95F30FF"),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test')
```


## Analyzing TAM_SPP1 levels in TCGA patients with or without EGFR amplification
```{r cnvEGFR-TAM_SPP1}
ggviolin(EGFR_CNV_riskscore, x = 'cnvlevel', y = '`TAM-SPP1`', fill = 'cnvlevel',
         palette = c("#639CA4FF","#BE7245FF"),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test')


ggviolin(EGFR_CNV_riskscore, x = 'cnvlevel', y = '`TAM_SPP1_sig_score`', fill = 'cnvlevel',
         palette = c("#639CA4FF","#BE7245FF"),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test')

```


## Analyzing TAM_CCL3 levels in TCGA patients with or without EGFR amplification
```{r cnvEGFR-TAM_CCL3}
ggviolin(EGFR_CNV_riskscore, x = 'cnvlevel', y = '`TAM-CCL3`', fill = 'cnvlevel',
         palette = c("#639CA4FF","#BE7245FF"),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test')


ggviolin(EGFR_CNV_riskscore, x = 'cnvlevel', y = '`TAM_CCL3_sig_score`', fill = 'cnvlevel',
         palette = c("#639CA4FF","#BE7245FF"),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test')

```


## Comparing risk scores in TCGA patients with or without EGFR amplification
```{r cnvEGFR-riskscore}
ggviolin(EGFR_CNV_riskscore, x = 'cnvlevel', y = 'riskscore', fill = 'cnvlevel',
         palette = c("#F8A02EFF","#9BB655FF"),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test')
```


## Comparing SPP1+TAM in TCGA patients with or without EGFR amplification
```{r}
x2 <- EGFR_CNV_riskscore[EGFR_CNV_riskscore$IDH.status %in% c("WT","Mutant"),]
x2$`TAM_SPP1_level`<-ifelse(x2$`TAM-SPP1`> median(x2$`TAM-SPP1`), "high","low")
```


## Creating sankey plots
```{r}
data_sankey = x2%>%
  dplyr::select(cnvlevel,TAM_SPP1_level)%>%
  mutate(cnvlevel=as.factor(cnvlevel),TAM_SPP1_level=as.factor(TAM_SPP1_level))%>%
  make_long(cnvlevel,TAM_SPP1_level)
str(data_sankey)

d_palettes <- palettes_d_names
col_sankey1 <- paletteer_d("colRoz::flavolineata",n=length(unique(data_sankey$node)))
```


```{r sankey-CNV-SPP1TAM}
ggplot(data_sankey, aes(x = x, next_x = next_x, node = node,
                        next_node = next_node, fill = factor(node), label = node)) +
  geom_alluvial(flow.alpha = .5,node.color = "gray30") +
  geom_alluvial_text(size = 0, color = "black") +
  scale_fill_manual(values =col_sankey1 )+
  theme_minimal(base_size = 12)+
  labs(x = NULL,fill='level') +
  ggtitle("CNV-SPP1TAM")+
  theme(legend.position = "right",
        plot.title = element_text(hjust = .5)) +
  theme(axis.text.y = element_blank())+
  scale_y_continuous(expand = c(0,0))
```


```{r}
data_sankey = x2%>%
  dplyr::select(cnvlevel,IDH.status,precoxlevel)%>%
  mutate(cnvlevel=as.factor(cnvlevel),status1=as.factor(IDH.status),precoxlevel=as.factor(precoxlevel))%>%
  make_long(cnvlevel,IDH.status,precoxlevel)
str(data_sankey)

d_palettes <- palettes_d_names 
col_sankey1<-paletteer_d("ggthemes::excel_Median",n=length(unique(data_sankey$node))) 
```


```{r sankey-IDH-CNV-riskscore}
ggplot(data_sankey, aes(x = x, next_x = next_x, node = node, 
                        next_node = next_node, fill = factor(node), label = node)) +
  geom_alluvial(flow.alpha = .5,node.color = "gray30") +
  geom_alluvial_text(size = 0, color = "black") +
  scale_fill_manual(values =col_sankey1 )+
  theme_minimal(base_size = 12)+
  labs(x = NULL,fill='level') +
  ggtitle("IDH-CNV-precox")+
  theme(legend.position = "right",
        plot.title = element_text(hjust = .5)) +
  theme(axis.text.y = element_blank())+
  scale_y_continuous(expand = c(0,0))
```


## Performing survival analysis
```{r}
x4<-EGFR_CNV_riskscore[EGFR_CNV_riskscore$cnvlevel==c("amplification"),]
fit<- survfit(Surv(OS, status) ~ `TAM_SPP1_level`, data = x4)
```

```{r survival_CNVgain_SPP1TAM1}
ggsurvplot(fit,pval = TRUE, palette = c("#E0C5BEFF","#FDD41AFF"),conf.int = TRUE,risk.table = TRUE, xlim = c(0, 3000),break.x.by=1000)

```

```{r}
x5<-EGFR_CNV_riskscore[EGFR_CNV_riskscore$IDH.status==c("Mutant"),]
x6<-x5[x5$cnvlevel==c("normal"),]
fit<- survfit(Surv(OS, status) ~ precoxlevel, data = x6)

```

```{r survival_IDHmut_normal_riskscore}
ggsurvplot(fit,pval = TRUE, palette = c("#DD8047FF","#A5AB81FF"),conf.int = TRUE,risk.table = TRUE, xlim = c(0, 3000),break.x.by=1000)

```

