---
title: "figure6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(randomForest) 
library(tidyverse) 
library(ggpubr) 
library(ROCR)
library(survival)
library(survminer)
library(glmnet)
library(survivalROC)
library(ggplot2)
library(Hmisc)
library(timeROC)
library(IOBR) 
library(ggsankey)
library(vioplot)
library(ggExtra)
```

## Performing univariate Cox regression analysis
```{r}
sig_gene <- read.table("./exp_tpm_sig_gene.csv", header=T, sep=",", check.names=F, row.names = 1)
cli <- read.table("./fig6_cli.csv", header=T, sep=",", check.names=F, row.names = 1)

cox <- apply(
  sig_gene,1,function(x){
    cli$genes <- as.numeric(x)
    cox_genes <- coxph(Surv(OS, status) ~ genes, data = cli)
    beta <- coef(cox_genes) 
    se <- sqrt(diag(vcov(cox_genes))) 
    HR <- exp(beta) 
    HRse <- HR * se
    cox_need <- round(cbind(coef = beta,
                            se = se,
                            z = beta/se,
                            p = 1 - pchisq((beta/se)^2, 1),
                            HR = HR, 
                            HRse = HRse,
                            HRz = (HR - 1) / HRse,
                            HRp = 1 - pchisq(((HR - 1)/HRse)^2, 1),
                            ower_95 = exp(beta - qnorm(.975, 0, 1) * se),
                            upper_95 = exp(beta + qnorm(.975, 0, 1) * se)), 3)
    return(cox_need["genes",])
  }
)

genes_cox <- t(cox)
genes_cox_sig <- genes_cox[genes_cox[,4]<0.05,]

x <- t(sig_gene[rownames(genes_cox_sig),])
y <- data.matrix(Surv(time = cli$OS,event = cli$status))

y<-y[-154,]
y<-y[-296,]
x<-x[-which(rownames(x) %in% c("TCGA-P5-A5EU","TCGA-P5-A5ET")),]

fit <- glmnet(x, y, family = 'cox', type.measure = "deviance", nfolds = 10)
```

## Conducting LASSO Cox regression analysis
```{r}
set.seed(123)
lasso_fit <- cv.glmnet(x, y, family = 'cox', type.measure = 'deviance', nfolds = 10)
lambda.min <- lasso_fit$lambda.min
model_lasso_min <- glmnet(x, y, family = 'cox', type.measure = 'deviance', nfolds = 10, lambda = 0.08)
gene_min <- rownames(model_lasso_min$beta)[as.numeric(model_lasso_min$beta)!=0]
```

## Conducting multivariate Cox regression analysis
```{r}
variable <- c("APOBEC3C","EMP3","IGF2BP2","TGIF1")
exp_4gene <- t(sig_gene[variable,])
exp_4gene_cli <- cbind(cli,exp_4gene)
exp_4gene_cli$status <- as.numeric(exp_4gene_cli$status)

model_4gene <- coxph(Surv(OS, status) ~ APOBEC3C+EMP3+IGF2BP2+TGIF1, data = exp_4gene_cli)

pre_cox <- predict(model_4gene,newdata = exp_4gene_cli)
with(exp_4gene_cli,rcorr.cens(pre_cox*(-1),Surv(OS, status),outx = T))
 
coef <- coef(model_4gene) 
df_ROC <- cbind(pre_cox,exp_4gene_cli)


Survival_ROC_input<-df_ROC
survival_ROC<-survivalROC(Stime=df_ROC$OS, 
                          status=df_ROC$status, 
                          marker=df_ROC$pre_cox, 
                          predict.time=1825, 
                          method="KM")

fit <- survfit(Surv(OS, status) ~ precoxlevel, data = df_ROC)
```



```{r risk_score_survival}
ggsurvplot(fit,pval = TRUE, palette = c("#F8A02EFF", "#9BB655FF"),conf.int = TRUE,risk.table = TRUE,xlim = c(0, 3000),break.x.by = 1000)
```

## Generating the receiver operating characteristic curve
```{r}
col <- c("#0073C2FF","firebrick1","orange") 
tt <- timeROC(df_ROC$OS,df_ROC$status,df_ROC$pre_cox,
              cause = 1,weighting = 'marginal',
              times = seq(365,1825,365),ROC = T,iid = T)
```

```{r 4gene-timeROC}
plot(tt,time=365 ,col=col[1],title=FALSE,lwd=1.5)
plot(tt,time=1095,col=col[2],add=TRUE,title=FALSE,lwd=1.5)
plot(tt,time=1825,col=col[3],add=TRUE,title=FALSE,lwd=1.5)
id <- c(paste0("1-Year AUC = ",round(tt$AUC[1],3)),
        paste0("3-Year AUC = ",round(tt$AUC[3],3)),
        paste0("5-Year AUC = ",round(tt$AUC[5],3)))
legend("bottomright",id,fill=col[1:3],bty="o",cex=1,border = NA)
abline(0,1,lty=2,lwd=0.5)
```



```{r}
model_multi2 <- coxph(Surv(OS, status) ~ `IDH.riskscore`, data =  cli) 
```

```{r forest_plot}
ggforest(model_multi2,data = cli,main = "Hazard ratio", cpositions = c(0.02, 0.16, 0.38),fontsize = 1.5,refLabel = "1", noDigits = 3) 
```

## Analyzing TAM subsets in high/low risk groups
```{r}
my_comparisons <- list( c("high", "low"))
p1 <- ggviolin(cli, x = 'precoxlevel', y = '`TAM-SPP1`', fill = 'precoxlevel',
               palette = c("#9BB655FF","#F8A02EFF"),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test') 

p2 <- ggviolin(cli, x = 'precoxlevel', y = '`TAM-AIF1`', fill = 'precoxlevel',
               palette = c("#9BB655FF","#F8A02EFF"),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test')

p3 <- ggviolin(cli, x = 'precoxlevel', y = '`TAM-CCL3`', fill = 'precoxlevel',
               palette = c("#9BB655FF","#F8A02EFF"),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test')

```





```{r riskscore_TAM_ciber}
p1
p2
p3
```

## Calculating score for signature genes
```{r}
my_signature<-list("res0.4_C0markers" = c("CCL3","CCL4","IL1B","CCL3L1","CCL4L2"),
                   "res0.4_C3markers" = c("C1QB","C1QA","HLA-DRB1","AIF1","CD74"),
                   "res0.4_C6markers" = c("SPP1","FTL","APOC1","S100A11","APOC2"),
                   "Immune_Checkpoint" = c("CD274","PDCD1LG2","CTLA4","PDCD1","LAG3","TIGIT"))

gene_exp_tpms <- read.table("./gene_exp_tpms.csv", header=T, sep=",", check.names=F, row.names = 1)

IOBRscore<-calculate_sig_score(eset = gene_exp_tpms,
                               signature = my_signature,
                               method = "pca",
                               mini_gene_count = 3,adjust_eset=TRUE)
tidy_cli <- read.table("./tidy_cli.csv", header=T, sep=",", check.names=F) 
```

## Generating waterfall plot
```{r waterfall}
ggplot(tidy_cli)+
  geom_bar(aes(reorder(Patient,TIDE),TIDE,fill=riskscorelevel),stat="identity",width=0.8) +
  labs(x="riskscorelevel",y="TIDE_score") +
  scale_fill_manual(values = c("#E7B800", "#2E9FDF")) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title = element_blank(),legend.position = c(0.6,0.9))
```


```{r}
my_comparisons <- list( c("High", "Low")) 
p1 <- ggviolin(tidy_cli, x = 'riskscorelevel',y = 'TIDE', fill = 'riskscorelevel',palette = c('#DE6DA7FF','#45837FFF'),add = 'boxplot',add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test')

p2 <- ggviolin(tidy_cli, x = 'riskscorelevel',y = 'Dysfunction', fill = 'riskscorelevel',palette = c('#DE6DA7FF','#45837FFF'),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test')

p3 <- ggviolin(tidy_cli, x = 'riskscorelevel',y = 'Exclusion', fill = 'riskscorelevel',palette = c('#DE6DA7FF','#45837FFF'),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test')
```

```{r riskscorelevel-TIDE}
p1
p2
p3
```


```{r}
rownames(IOBRscore) <- substr(IOBRscore$ID,1,12) 
sameSample=intersect(row.names(IOBRscore), row.names(df_ROC))
rt=cbind(IOBRscore[sameSample,,drop=F], df_ROC[sameSample,,drop=F])
rt <- rt[,c(2:6)]
```


```{r risk_score_cor}
outTab=matrix(ncol=3,nrow=1)
for(i in colnames(rt)[1:1]){
  x=as.numeric(rt[,"pre_cox"])
  y=as.numeric(rt[,i])
  cor=cor.test(x, y, method="spearman")
  outVector=cbind(Cell=i, cor=cor$estimate, pvalue=cor$p.value)
  outTab=rbind(outTab,outVector)
  if(cor$p.value<0.05){
    df1=as.data.frame(cbind(x,y))
    p1=ggplot(df1, aes(x, y)) + 
      xlab(paste0("risk_score")) + ylab("TAM_CCL3_sig_score")+
      geom_point() + geom_smooth(method="lm",formula = y ~ x) + theme_bw()+
      stat_cor(method = 'spearman', aes(x =x, y =y))
    p2=ggMarginal(p1, type="density", xparams=list(fill = "orange"), yparams=list(fill = "#DE6DA7FF"))
    print(p2)
  }
}
```

```{r risk_score_cor3}
outTab=matrix(ncol=3,nrow=1)
for(i in colnames(rt)[3:3]){
  x=as.numeric(rt[,"pre_cox"])
  y=as.numeric(rt[,i])
  cor=cor.test(x, y, method="spearman")
  outVector=cbind(Cell=i, cor=cor$estimate, pvalue=cor$p.value)
  outTab=rbind(outTab,outVector)
  if(cor$p.value<0.05){
    df1=as.data.frame(cbind(x,y))
    p1=ggplot(df1, aes(x, y)) + 
      xlab(paste0("risk_score")) + ylab("TAM_SPP1_sig_score")+
      geom_point() + geom_smooth(method="lm",formula = y ~ x) + theme_bw()+
      stat_cor(method = 'spearman', aes(x =x, y =y))
    p2=ggMarginal(p1, type="density", xparams=list(fill = "orange"), yparams=list(fill = "#45837FFF"))
    print(p2)
  }
}
```



```{r risk_score_cor2}
outTab=matrix(ncol=3,nrow=1)
for(i in colnames(rt)[2:2]){
  x=as.numeric(rt[,"pre_cox"])
  y=as.numeric(rt[,i])
  cor=cor.test(x, y, method="spearman")
  outVector=cbind(Cell=i, cor=cor$estimate, pvalue=cor$p.value)
  outTab=rbind(outTab,outVector)
  if(cor$p.value<0.05){
    df1=as.data.frame(cbind(x,y))
    p1=ggplot(df1, aes(x, y)) + 
      xlab(paste0("risk_score")) + ylab("TAM_AIF1_sig_score")+
      geom_point() + geom_smooth(method="lm",formula = y ~ x) + theme_bw()+
      stat_cor(method = 'spearman', aes(x =x, y =y))
    p2=ggMarginal(p1, type="density", xparams=list(fill = "orange"), yparams=list(fill = "darkorange"))
    print(p2)
  }
}
```


## Validating findings on CGGA datasets
```{r CGGA_score_survival}
df_cgga <- read.csv("df_cgga.csv", row.names = 1)
fit<- survfit(Surv(OS, status) ~ test, data = df_cgga)

ggsurvplot(fit,pval = TRUE,palette = c("#EA879CFF", "#1BB6AFFF"),conf.int = TRUE,risk.table = TRUE, xlim = c(0, 3000),break.x.by=1000)

```


## Plotting ROC curves at multiple time points
```{r}
col <- c("#0073C2FF","firebrick1","orange")
tt <- timeROC(df_cgga$OS,df_cgga$status,df_cgga$riskscore,
              cause = 1,weighting = 'marginal',
              times = seq(365,1825,365),ROC = T,iid = T)

tt$AUC
```


```{r CGGA-4gene-timeROC}
plot(tt,time=365 ,col=col[1],title=FALSE,lwd=1.5)
plot(tt,time=1095,col=col[2],add=TRUE,title=FALSE,lwd=1.5)
plot(tt,time=1825,col=col[3],add=TRUE,title=FALSE,lwd=1.5)
id <- c(paste0("1-Year AUC = ",round(tt$AUC[1],3)),
        paste0("3-Year AUC = ",round(tt$AUC[3],3)),
        paste0("5-Year AUC = ",round(tt$AUC[5],3)))
legend("bottomright",id,fill=col[1:3],bty="o",cex=1,border = NA)
abline(0,1,lty=2,lwd=0.5)
```


```{r}
df_cgga$TIDElevel=ifelse(df_cgga[,"TIDE"]>median(df_cgga[,"TIDE"]), "TIDE-High", "TIDE-Low")
df_cgga$TIDElevel<-as.factor(df_cgga$TIDElevel)
```


## Comparing TIDE Scores in High/Low Risk Groups
```{r CGGA-riskscore-TIDE}
my_comparisons <- list( c("High", "Low")) 
ggviolin(df_cgga, x = 'riskscorelevel',y = 'TIDE', fill = 'riskscorelevel',palette = c("#DE6DA7FF","#45837FFF"),add = 'boxplot',add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test') 

ggviolin(df_cgga, x = 'riskscorelevel',y = 'Exclusion', fill = 'riskscorelevel',palette = c("#DE6DA7FF","#45837FFF"),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test')

```


## Generating a TIDE waterfall plot
```{r riskscore-TIDE-waterfall}
ggplot(df_cgga)+
  geom_bar(aes(reorder(Patient,TIDE),TIDE,fill=riskscorelevel),stat="identity",width=0.8) +
  labs(x="riskscore",y="TIDE_score") +
  scale_fill_manual(values = c("#E7B800", "#2E9FDF"))+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title = element_blank(),legend.position = c(0.6,0.9))

```


```{r CGGA_riskscore_TAMs}
my_comparisons <- list( c("High", "Low"))
ggviolin(df_cgga, x = 'riskscorelevel', y = 'ciber_TAM_SPP1', fill = 'riskscorelevel',
               palette = c("#1BB6AFFF","#EA879CFF"),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 't.test') 


ggviolin(df_cgga, x = 'riskscorelevel', y = 'ciber_TAM_AIF1', fill = 'riskscorelevel',
               palette = c("#1BB6AFFF","#EA879CFF"),add = 'boxplot', add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif", bracket.size=0.5, tip.length = 0.02, method = 'wilcox.test')

```



